%\newpage
\begin{center}
{\LARGE \bf LIST OF PUBLICATIONS}
\end{center}
\vspace{0.2in}

\renewcommand{\labelenumi}{\arabic{enumi}.}   % Added on 09/08/2008
\renewcommand{\theenumii}{\arabic{enumii}}
\renewcommand{\labelenumii}{\theenumii}
%\noindent{\large \bf \textit{LIST OF PUBLICATIONS}} \\
%\vspace{0.2cm}
\noindent{\bf{Journal Publications}}
\begin{enumerate}
	\item {\label{IET2016}} {\bf Sunil Kumar}, M.K. Bhuyan and B. K. Chakraborty, ``Extraction of informative regions of a face for facial expression recognition", \emph{IET Computer Vision}, vol. 10, no. 6, pp. 567-576, 2016.
	
	\item {\label{JMUI2016}} {\bf Sunil Kumar}, M.K. Bhuyan and B. K. Chakraborty, ``Extraction of Texture and Geometrical Features from Informative Facial Regions for Sign Language Recognition", \emph{Journal on Multimodal User Interfaces, Springer}, pp. 1-13, 2017.
	
\end{enumerate}
\noindent{\bf Manuscripts under Review}
\begin{enumerate}
	\item {\label{CSVT2017}} {\bf Sunil Kumar} and M.K. Bhuyan ``Multilevel Uncorrelated Discriminative Shared Gaussian Process for Multiview Facial Expression Recognition", \emph{IEEE Transactions on Circuits Systems Video Technology}, {\bfseries{(Date of communication : 03-Nov-2016)}}.
	
	\item  {\label{AC2017}} {\bf Sunil Kumar} and M.K. Bhuyan ``Hierarchical Uncorrelated Multiview Discriminant Locality Preserving Projection for Multiview Facial Expression Recognition", \emph{IET Biometrics}, {\bfseries{(To be communicated)}}.
	
\end{enumerate}
\noindent{\bf {Conference Publications}}
\begin{enumerate}
	
	\item {\label{ICVGIP2016}} {\bf Sunil Kumar}, M.K. Bhuyan, and Biplab Ketan Chakraborty, ``Uncorrelated multiview discriminant locality preserving projection analysis for multiview facial expression recognition", \emph{Proceedings of Tenth Indian Conference on Computer Vision, Graphics and Image Processing (ICVGIP 2016).} ACM, 2016.
	
	\item {\label{NCC2016}} {\bf Sunil Kumar}, M.K. Bhuyan and B. K. Chakraborty, ``An efficient face model for facial expression recognition", \emph{Proceedings of Twenty Second National Conference on Communication (NCC 2016), Guwahati.} pp. 1-6, 2016.
	
	\item {\label{RAICS2015}} {\bf Sunil Kumar}, and M.K. Bhuyan. ``Neutral expression modeling in feature domain for facial expression recognition", \emph{Proceedings of IEEE conference on Recent Advances in Intelligent Computational Systems (RAICS 2015)}, pp. 224-228, 2015.
	
\end{enumerate}

%\noindent{\bf{Refereed Journals:}}
%\begin{enumerate}
%\item {\label{IET2016}} Sunil Kumar, M. K. Bhuyan and B. K. Chakraborty, ``Extraction of informative regions of a face for facial expression recognition", \emph{in IET Computer Vision}, vol. 10, no. 6, pp. 567-576, 9 2016.
%\end{enumerate}
%
%\item  {\label{JMUI2016}} Sunil Kumar, M. K. Bhuyan and B. K. Chakraborty, ``Extraction of Texture and Geometrical Features from Informative Facial Regions for Sign Language Recognition", \emph{Journal on Multimodal User Interfaces, Springer}, 2016.
%
%\noindent{\bf Manuscripts under Review}
%\begin{enumerate}
%\item {\label{CSVT2017}} Sunil Kumar, M. K. Bhuyan and Akhil Babu Manam, ``Multilevel Uncorrelated Discriminative Shared Gaussian Process for Multiview Facial Expression Recognition", \emph{IEEE Transactions on Circuits Systems Video Technology}, 2016.
%	
%\item {\label{AC2017}} Sunil Kumar, M. K. Bhuyan, B. K. Chakraborty and Yuji Iwahori ``Hierarchical Uncorrelated Multiview Discriminant Locality Preserving Projection for Multiview Facial Expression Recognition", \emph{IEEE Transactions on Affective Computing}, 2016.
%	
%\end{enumerate}
%\noindent{\bf {Refereed Conferences:}}
%\begin{enumerate}
%
%\item {\label{ICVGIP2016}} Sunil Kumar, M. K. Bhuyan, and Biplab Ketan Chakraborty, ``Uncorrelated multiview discriminant locality preserving projection analysis for multiview facial expression recognition", \emph{Proceedings of the Tenth Indian Conference on Computer Vision, Graphics and Image Processing.} ACM, 2016.
%
%\item {\label{NCC2016}} Sunil Kumar, M. K. Bhuyan and B. K. Chakraborty, ``An efficient face model for facial expression recognition", \emph{2016 Twenty Second National Conference on Communication (NCC), Guwahati.} pp. 1-6, 2016.
%
%\item {\label{RAICS2015}} Sunil Kumar, and M. K. Bhuyan. ``Neutral expression modeling in feature domain for facial expression recognition", \emph{Intelligent Computational Systems (RAICS), 2015 IEEE Recent Advances in. IEEE}, 2015.
%
%\end{enumerate}
